---
title: "p8105_hw2_zl2978"
author: Zhiye Lou
output: github_document
---
I will first load the library.
```{r}
library(tidyverse)
library(readxl)
```

## Problem 1
I will first set the path to the data.
```{r}
path_to_data_TW = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Then I will read the data and clean the data.
```{r}
trashwheel_df = read_xlsx(
    path = path_to_data_TW,
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
janitor::clean_names() %>% 
drop_na(dumpster) %>% 
mutate(
    sports_balls = round (sports_balls),
    sports_balls = as.integer (sports_balls)
)
```

Then I will read and clean precipitation data for 2017 and 2018.
```{r}
ppt_2017 = read_xlsx(
    path = path_to_data_TW,
    sheet = "2017 Precipitation",
    skip = 1) %>% 
janitor::clean_names() %>% 
drop_na(month) %>% 
mutate(year = 2017) %>% 
relocate(year)

ppt_2018 = read_xlsx(
    path = path_to_data_TW,
    sheet = "2018 Precipitation",
    skip = 1) %>% 
janitor::clean_names() %>% 
drop_na(month) %>% 
mutate(year = 2018) %>% 
relocate(year)
```

Next, I will combine precipitation datasets and convert month to a character variable.
```{r}
month_df = 
    tibble(month = 1:12,
           month_name = month.name)
precip_df = 
	  bind_rows(ppt_2017, ppt_2018)
precip_df =
  	left_join(precip_df, month_df, by = "month") %>% 
select(-month)
```

This dataset reports data from the Mr. Trashwheel trash collector in Baltimore, Maryland. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. For the data of Trash Wheel, there are a total of `r nrow(trashwheel_df)` rows. The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.Additional data sheets include month precipitation data in year 2017 and year 2018. The dataset contains information on year, month, and total precipitation in each month. There are a total of `r nrow(precip_df)` rows. The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. The total precipitation in 2017 was `r precip_df %>% filter(year == 2017) %>% pull(total) %>% sum()` inches. 

##Problem 2
I will read and clean the data.
```{r}
NYC_Transit_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_data.csv") %>% 
janitor::clean_names() %>% 
select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>% 
mutate(entry = if_else(entry == "YES", T, F)) 
```
This dataset contains information on line, station name, specific location (longitude, latitude), route information, if there is an entry,if there is a vending, the entrance type, and if there is an APA compliance of each subway station in NYC. 
I have clean the names of this dataset, drop several variabes that we are not interested in and mutate the entry variable to a logical variable.
There are a total of `r nrow(NYC_Transit_data)` rows and a total of `r ncol(NYC_Transit_data)` columns.
I do not think the data is tidy enough. For the route variables, I think we may combine them into a single variable. For example, for the first row, the value of route could be "RNNNNNNNNNN" where N stands for NA, and this will reduce 10 columns for our final data frame. In addition, since we dropped the variable "corner", several rows appear to be totally same and the difference between different rows which represent the same stations is relatively meaningless.For example, the reader of this dataset will not understand why row 4 and row 5 exist and the same time.

There are `r nrow(distinct(NYC_Transit_data, line, station_name))` distinct stations, and `r filter(NYC_Transit_data, ada == T) %>% distinct(line,station_name) %>% nrow()` stations are ADA compliant. The proportion of station entrances / exits without vending allow entrance is `r filter(NYC_Transit_data, entry == T) %>% filter(vending == "NO") %>% nrow()/nrow(filter(NYC_Transit_data, vending == "NO"))`

Then I will reformat data so that route number and route name are distinct variables.
```{r}
NYC_reformat_transit = 
NYC_Transit_data %>% 
mutate(
  route8 = as.character(route8),
  route9 = as.character(route9),
  route10 = as.character(route10),
  route11 = as.character(route11)) %>% 
pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number" 
) %>% 
mutate(route_name = substr(route_name,6,8)) 
```
There are `r filter(NYC_reformat_transit, route_number == "A") %>% distinct(line, station_name) %>% nrow()` distinct stations serve the A train, and `r filter(NYC_reformat_transit, route_number == "A") %>% filter(ada == T) %>% distinct(line, station_name) %>%  nrow()` are ADA compliant.

##Problem 3
I will first read and tidy the data.
```{r}
pols_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
janitor::clean_names() %>% 
separate(mon, into = c("year","month","day"), sep = "-") %>%
mutate(month = as.integer(month)) %>% 
left_join(month_df, by = "month") %>% 
mutate(month = month_name) %>% 
select(-month_name) %>% 
mutate(president = recode(prez_gop, "1" = "gop", "0" = "dem")) %>% 
select(-prez_dem,-prez_gop,-day) %>% 
relocate(year,month,president) 
```
I did not recode the prez_gop = 2 and it changes to NA automatically.

Then I will clean the data in snp.csv.
```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
janitor::clean_names() %>% 
separate(date, into = c("month","day","year"), sep = "/") %>%
mutate(month = as.integer(month)) %>% 
left_join(month_df, by = "month") %>% 
mutate(month = month_name) %>% 
select(-month_name,-day) %>% 
relocate(year,month) 
```
I drop the day variable so that it can be merged with the other two data and easier to understand.

Then I will tidy the unemployment data.
```{r}
monthabb_df = 
    tibble(month = tolower(month.abb),
           month_name = month.name)
unemp_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
janitor::clean_names() %>% 
pivot_longer(
  jan:dec,
  names_to = "month",
  values_to = "unemployment_rate"
) %>% 
left_join(monthabb_df, by = "month") %>% 
mutate(month = month_name) %>% 
select(-month_name)
```

Then I will join the datasets.
```{r}
merge_df =
  merge(pols_df,snp_df) %>% 
merge(unemp_df)
```

Until now, I have 4 datasets in total. The first one is pols_df which contains the information about the year, month, president type, the number of republican and democratic governors, the number of republican and democratic senators and the number of republican and democratic representatives on a given date. There are `r nrow(pols_df)` rows and `r ncol(pols_df)` columns. The range of the year is from `r min(pull(pols_df,year))` to `r max(pull(pols_df,year))`, one important variable in this data frame is the president which represents if the president is democratic or republican.

The second one is snp.csv which contains the information about the year, month and the closing values of the S&P stock index on the associated date. The variable close represents the closing values of the stock index. There are `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. The range of the year is from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. Besides, I drop the day variable for better merging with the other two data frame.

The third one is unemp_df which contains the information about the year, month and the unemployment rate at that time. The variable unemployment_rate represents the unemployment rate. There are `r nrow(unemp_df)` rows and `r ncol(unemp_df)` columns. The range of the year is from `r min(pull(unemp_df,year))` to `r max(pull(unemp_df,year))`.

The last one is the dataset that merge all three datasets and information together. I have clean the names of all these three datasets. I have convert all the data variable in these three dataset to a same format which are year and month, and the month value are expressed as words. Besides, I have relocate the year and month as the first two columns.

The last dataset contains the information about the year, month, president type, the number of republican and democratic governors, the number of republican and democratic senators and the number of republican and democratic representatives on a given date, closing values of the stock index, and the unemployment rate. Key variables are president, unempolyment_rate and close. There are `r nrow(merge_df)` rows and `r ncol(merge_df)` columns. The range of the year is from `r min(pull(merge_df,year))` to `r max(pull(merge_df,year))`.
